\documentclass[11pt,a4paper]{article}

% Packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{cite}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{booktabs}
\usepackage{geometry}
\geometry{margin=1in}

% Title
\title{Rotation Trick for Vector Quantized Variational Autoencoders}
\author{AI Researcher}
\date{\today}

\begin{document}

\maketitle

% Abstract
\input{abstract}

% Introduction
\input{introduction}

% Related Work
\input{related_work}

% Methodology
\input{methodology}

% Experiments
\input{experiments}

% Conclusion
\input{conclusion}

% References
\begin{thebibliography}{10}
    \bibitem{vqvae} Oord, A. v. d., Dieleman, S., \& Korzeniowski, K. (2017). Neural Discrete Representation Learning. \textit{NeurIPS}.
    \bibitem{botta2020} Botta, E., et al. (2020). Gaussian Mixture Variational Autoencoders. \textit{ICLR}.
    \bibitem{gumbelsoftmax} Jang, E., Gu, S., \& Poole, B. (2017). Categorical Reparameterization with Gumbel-Softmax. \textit{ICLR}.
    \bibitem{kingma2013} Kingma, D. P., \& Welling, M. (2013). Auto-Encoding Variational Bayes. \textit{ICLR}.
    \bibitem{latdiff} Hundt, C., et al. (2021). High-Resolution Image Synthesis with Latent Diffusion Models. \textit{CVPR}.
    \bibitem{compvis2021} CompVis. (2021). Latent Diffusion: High-Quality Image Generation via Latent Variable Models. \textit{NeurIPS}.
    \bibitem{ho2020} Ho, J., et al. (2020). Denoising Diffusion Probabilistic Models. \textit{NeurIPS}.
    \bibitem{classifierfree} Ho, J., \& Salimans, T. (2022). Classifier-Free Guidance for Efficient Image Generation. \textit{NeurIPS}.
    \bibitem{chen2021} Chen, M., et al. (2021). Text-guided image synthesis with generative adversarial networks. \textit{ACM Transactions on Graphics}.
\end{thebibliography}

\end{document}
